---
title: "code_challlenge_bella"
format: html
editor: source
---

# Code Challenge:

##1. Load Data Into Environment

```{r}
bps_aoi_attributes <- read.csv("code_challenge/input_data/bps_aoi_attributes.csv")

bps_model_number_names <- read.csv("code_challenge/input_data/bps_model_number_name.csv")

combine_raw <- read.csv("code_challenge/input_data/combine_raw.csv")

LF16_BPS_200 <- read.csv("code_challenge/input_data/LF16_BPS_200.csv")

ref_con_modified <- read.csv("code_challenge/input_data/ref_con_modified.csv")

scls_aoi_attributes <- read.csv("code_challenge/input_data/scls_aoi_attributes.csv")

```

###Practice chart

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)

ggplot(scls_aoi_attributes, aes(x=REL_PERCENT, y=LABEL)) + 
  geom_bar(stat = "identity")
```

##2. Merge & Filter for past data
```{r}
merged_modelcode <- merge(bps_model_number_names, ref_con_modified, by = "Model_Code")

merged_mc_pivot <- pivot_longer(merged_modelcode,
                          cols = c(A:E, Agriculture:UE))

filtered_pivot <- filter(merged_mc_pivot, Model_Code == "13040_32_43_44_49" |  Model_Code == "15070_44" | Model_Code == "13670_32_44")

colnames(filtered_pivot)[colnames(filtered_pivot) == "value"] <- "Rel_Percent"

```

### Practice Chart (Past only):
```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)

ggplot(filtered_pivot, aes(x=Rel_Percent, y=name)) + 
  geom_bar(stat = "identity")

```

#Merging for Present Data
```{r}
colnames(scls_aoi_attributes)[colnames(scls_aoi_attributes) == "VALUE"] <- "Var2"

merged_variance <- left_join(combine_raw, scls_aoi_attributes %>% dplyr::select(2, 4), by = "Var2")

colnames(bps_aoi_attributes)[colnames(bps_aoi_attributes) == "VALUE"] <- "Var1" #I think I did this because I thought it would connect merged_variance/ help in calculations but I actually don't think it will

colnames(LF16_BPS_200)[colnames(LF16_BPS_200) == "VALUE"] <- "Var1" #similarly wanted to change to Var1 to possibly filter/merge with merged_variance

mergedv2 <- left_join(merged_variance, LF16_BPS_200 %>% dplyr::select(1:5), by = "Var1")

# calculate current sclass percents
combine <- mergedv2 %>%
  group_by(Var1, BPS_MODEL) %>%
  mutate(total_count = sum(Freq)) %>%
  mutate(currentPercent = as.integer((Freq/total_count)*100)) %>%
  unite(model_label, c("BPS_MODEL", "LABEL"))

```

##Trying to filter?
```{r}

filtered_combine <- filter(combine, BPS_CODE == "13040" | BPS_CODE == "13670" | BPS_CODE == "15070")

filtered_combine$BPS_NAME <- factor(filtered_combine$BPS_NAME, levels = unique(filtered_combine$BPS_NAME))

# Reorder the dataset by the BPS_NAME column
reordered_combine <- arrange(filtered_combine, BPS_NAME)

```

>merged_total <- merge(filtered_aoi, merged_variance, by = "Var1")

>  I wish I had been more intentional in my naming because it made things more difficult leaving them with their original names, I would have to keep looking through the datasets to remember what I needed from them. 

#Put it together
```{r}
library(janitor)

past_and_present <- cross_join(filtered_pivot, reordered_combine) %>% 
  drop_na("Rel_Percent") %>% 
  mutate(currentPercent = as.numeric(currentPercent), currentPercent = ifelse(is.na(currentPercent), 0, currentPercent)) %>%
          mutate(total_count = as.numeric(total_count),  total_count = ifelse(is.na(total_count), 0, total_count)) %>%
          select(-c(BPS_CODE, ZONE)) %>%
                         select(c(Freq, 
                                  Var1, 
                                  Var2,
                                  BpS_Name,
                                  Model_Code,
                                  model_label,
                                  Rel_Percent,
                                  currentPercent,
                                  total_count)) %>%
                          rename(count = Freq,
                                 bps_value = Var1,
                                 scl_value = Var2,
                                 bps_name = BpS_Name) %>%
                          clean_names() 
```

## Make Chart
```{r}
# Load the tidyr library
library(tidyr)




```


#Reflection:

ii. From this challenge my comfortability with data wrangling has improved tremendously. At the start of this challenge I was still rusty/ not completely comfortable and having to ask chatgpt questions about inputting data and merging datasets. Now I feel that if we were given a second code challenge, with the information I've learned from this first, I might be able to work through things more efficiently.

iii. I think I did a good job at reflecting/ note taking as I went along and struggled. I made notes for myself so if I took a break and came back another day, or a few days later, I still knew what I had going on and needed to work with. I think if I were working with a partner on something like this that form of note taking would be helpful and make things easier for two people. Similarly, I could definitely improve on making comments within code, that's something I need to get in the habit of doing.

##Annotations:

> merge: merge is used to merge 2 datasets together by a specified column.

>pivot_longer: apart of "tidyr," used to shift datasets from a wide to long format.

>filter: used to filtered datasets by a specific attribute within a column.

>colnames: used to rename columns.

>left_join: used to join the left side of one dataset to all the matching attributes on the right of another dataset.

>group_by: used in conjunction with "mutate()" or other functions of the "dplyr" package,enabling you to group one or more variables in a dataset and manipulate them by their group.

>unite: concentrates values in R by tidying and uniting messy columns.
